---
title: "COVID 19 "
author: "VincentZ"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(hms)
library(lubridate)
library(ggplot2)
library(tidyverse)
```

```{r get_jhu_data, message = FALSE}
url_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"
file_names <- c("time_series_covid19_confirmed_global.csv","time_series_covid19_deaths_global.csv", "time_series_covid19_confirmed_US.csv", "time_series_covid19_deaths_US.csv" )
urls <- str_c(url_in, file_names)

global_cases <- read_csv(urls[1])
global_deaths <- read_csv(urls[2])
US_cases <- read_csv(urls[3])
US_deaths <- read_csv(urls[4])
```

```{r transformaing data}
global_cases <- global_cases %>% 
    pivot_longer(cols = -c('Province/State', 'Country/Region', 'Lat', 'Long'), names_to = "date", values_to = "cases") %>%
    select (-c(Lat,Long))

global_deaths <- global_deaths %>% 
    pivot_longer(cols = -c('Province/State', 'Country/Region', 'Lat', 'Long'), names_to = "date", values_to = "deaths") %>%
    select (-c(Lat,Long))

global <- global_cases %>%
  full_join(global_deaths)  %>%
  rename (Country_Region = 'Country/Region',
          Province_State = 'Province/State') %>%
  mutate(date =  mdy(date))

summary(global)


global <- global %>% filter (cases > 0)
summary(global)
```



```{r class_transformations}
US_cases <- US_cases %>%
  pivot_longer(cols = -(UID:Combined_Key),
               names_to  = "date",
               values_to = "cases") %>%
  select(Admin2:cases) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat, Long_))

US_cases

US_deaths <- US_deaths %>%
  pivot_longer(cols = -(UID:Population),
               names_to  = "date",
               values_to = "deaths") %>%
  select(Admin2:deaths) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat, Long_))

US_deaths
US <- US_cases %>%
  full_join(US_deaths)

global <- global %>%
  unite("Combined_Key",
        c(Province_State, Country_Region),
        sep = ", ",
        na.rm = TRUE,
        remove = FALSE)
global
```


```{r}
uid_lookup_url  <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"

uid <- read_csv(uid_lookup_url) %>%
  select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))

global <- global %>%
  left_join(uid, by = c("Province_State", "Country_Region")) %>%
  select(-c(UID, FIPS))

global

```



### I wanted to see the case per death relation so i made a scatterplot regarding global cases and deaths



```{r scatter_plot}
ggplot(global, aes(x = cases, y = deaths)) +
  geom_point(alpha = 0.6, color = "red") + 
  labs(
    title = "Global COVID-19 Cases and Deaths",
    x = "Global Cases",
    y = "Global Deaths"
  ) +
  theme_minimal()
```



### I then created a linear model that sees how the correlation of hte scatterplot above with cases and deaths are in the linear model.



```{r lm_model_cases_deaths}
ggplot(global, aes(x = cases, y = deaths)) +
  geom_point(alpha = 0.5, color = "blue") + 
  geom_smooth(method = "lm", color = "red", se = TRUE) + 
  labs(
    title = "Linear Model of Global COVID-19 Cases and Deaths",
    x = "Global Cases",
    y = "Global Deaths"
  ) +
  theme_minimal()

```



### Another visualization on the degree of the case. as there are way more cases the death counter is not as big in a larger scale



```{r time_series_cases}
ggplot(global, aes(x = date)) +
  geom_line(aes(y = cases, color = "Cases")) +
  geom_line(aes(y = deaths, color = "Deaths")) +
  labs(
    title = "Global COVID-19 Cases and Deaths",
    x = "Date",
    y = "Count",
    color = "Metric"
  ) +
  theme_minimal()

```



### I listed the top ten countries that had the most cases as well as deaths on the next one




```{r top_ten_countries_cases}
top_countries <- global %>%
  group_by(Country_Region) %>%
  summarise(total_cases = sum(cases), total_deaths = sum(deaths)) %>%
  arrange(desc(total_cases)) %>%
  slice(1:10)

ggplot(top_countries, aes(x = reorder(Country_Region, -total_cases), y = total_cases)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(
    title = "Top 10 Countries by Total COVID-19 Cases",
    x = "Country",
    y = "Total Cases"
  ) +
  coord_flip() +
  theme_minimal()

```


```{r top_ten_countries_deaths}
top_countries <- global %>%
  group_by(Country_Region) %>%
  summarise(total_cases = sum(cases), total_deaths = sum(deaths)) %>%
  arrange(desc(total_deaths)) %>%
  slice(1:10)

ggplot(top_countries, aes(x = reorder(Country_Region, -total_deaths), y = total_deaths)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Top 10 Countries by Total COVID-19 Deaths",
    x = "Country",
    y = "Total Deaths"
  ) +
  coord_flip() +
  theme_minimal()

```



