---
title: "NYPD Report"
author: "Vincent Zagala"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(hms)
library(lubridate)
library(ggplot2)
```
# Introduction
Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. In addition, information related to suspect and victim demographics is also included. This data can be used by the public to explore the nature of shooting/criminal activity.


### Including the libraries as well as importing the data from the website
```{r}
library(dplyr)
library(hms)
library(lubridate)

# Taking the Data from the Site and saving each column saved into a variable
nypd_cases <- read.csv(url("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"))
```
### Summarized data from the data set before any changes is made
```{r}
summary(nypd_cases)
```
### Renaming Columns for Easier Reference
```{r eval=FALSE}
nypd_cases_1 <- data.frame(
  INCIDENT_KEY = Incident_Key, 
  OCCUR_DATE = Occur_Date, 
  OCCUR_TIME = Occur_Time, 
  BORO = Location_Of_Incident, 
  PRECINCT = Precinct, 
  JURISDICTION_CODE = Jurisdiction_Code, 
  PERP_AGE_GROUP = Suspect_Age_Group,
  PERP_SEX = Suspect_Sex, 
  PERP_RACE = Suspect_Race, 
  VIC_AGE_GROUP = Victim_Age_Group, 
  VIC_SEX = Victim_Sex, 
  VIC_RACE = Victim_Race, 
  X_COORD_CD = X_Coord, 
  Y_COORD_CD = Y_Coord, 
  Latitude = Lat, 
  Longitude = Long
)

nypd_cases_1 <- nypd_cases %>%
    select(
        INCIDENT_KEY,
        OCCUR_DATE,
        OCCUR_TIME,
        BORO,
        PRECINCT,
        JURISDICTION_CODE,
        PERP_AGE_GROUP,
        PERP_SEX,
        PERP_RACE,
        VIC_AGE_GROUP,
        VIC_SEX,
        VIC_RACE,
        X_COORD_CD,
        Y_COORD_CD,
        Latitude,
        Longitude
    )
```

### I edited columns that needed to be converted into the appropriate data type such as Occur_Date converted into Date dataype and changed the murder flag from just characters into a boolean value based on true or false 


```{r}
nypd_cases_1 <- nypd_cases %>%
    rename(
        Incident_Key = INCIDENT_KEY, 
        Occur_Date = OCCUR_DATE, 
        Occur_Time = OCCUR_TIME, 
        Location_Of_Incident = BORO, 
        Precinct = PRECINCT, 
        Jurisdiction_Code = JURISDICTION_CODE, 
        Suspect_Age_Group = PERP_AGE_GROUP,
        Suspect_Sex = PERP_SEX, 
        Suspect_Race = PERP_RACE, 
        Victim_Age_Group = VIC_AGE_GROUP, 
        Victim_Sex = VIC_SEX, 
        Victim_Race = VIC_RACE, 
        X_Coord = X_COORD_CD, 
        Y_Coord = Y_COORD_CD, 
        Lat = Latitude, 
        Long = Longitude
    ) %>%
    mutate(
        Occur_Date = as.Date(Occur_Date, format = "%m/%d/%Y"),
        Occur_Time = hms::as_hms(Occur_Time),
        STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG == "Y"
    )
```   
    
### Bias

#### Possible biases in the data can arise from various sources. For example, if the data analyst is familiar with the area or has a background in criminal justice, their knowledge of criminal tendencies could introduce bias.

#### Reporting Bias: If crimes are not reported, they will not be included in the dataset. Conversely, if crimes are reported excessively, this can lead to an inaccurate representation of the number of cases for a given year or location.

#### Jurisdictional Bias: In the NYPD dataset, if multiple precincts report the same crime, it can result in duplicate entries. Alternatively, if two precincts are involved, each might assume the other will make the report, leading to no report being filed. This can result in underreporting of cases.
    
    
    
### Linear Model Analysis
    
#### The model I'm looking at is the Linear model in which I'm trying to find the number of cases per year
```{r}
nypd_cases_1 <- nypd_cases_1 %>%
mutate(Occur_Date = as.Date(Occur_Date, format = "%m/%d/%Y")) %>%
mutate(YEAR = format(Occur_Date, "%Y"))

cases_per_year <- nypd_cases_1 %>%
  group_by(YEAR) %>%
  summarise(CASE_COUNT = n())


lm_model <- lm(CASE_COUNT ~ as.numeric(YEAR), data = cases_per_year)


summary(lm_model)
```  


#### The linear model shows that the cases per year has been decreasing over time and that since theres a big decrease throughout the recent years. other factors may affect it that was not captured in the years.

#### I then made a linear graph regarding the actual case numbers per year with the predicted number of cases.


```{r}
cases_per_year <- cases_per_year %>%
  mutate(PREDICTED_CASE_COUNT = predict(lm_model, newdata = cases_per_year))

ggplot(cases_per_year, aes(x = as.numeric(YEAR))) +
  geom_line(aes(y = CASE_COUNT, color = "Actual Cases"), linewidth = 1.2) +
  geom_line(aes(y = PREDICTED_CASE_COUNT, color = "Predicted Cases"), 
            linetype = "dashed", linewidth = 1.2) +
  labs(
    title = "Actual vs Predicted NYPD Cases per Year",
    x = "Year",
    y = "Number of Cases",
    color = "Legend"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```
    
```{r}
# Calculate the residuals from the linear model
cases_per_year <- cases_per_year %>%
  mutate(RESIDUALS = residuals(lm_model))

ggplot(cases_per_year, aes(x = as.numeric(YEAR), y = RESIDUALS)) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residual Plot: Actual vs Predicted NYPD Cases per Year",
    x = "Year",
    y = "Residuals"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```



#### The residual plot provides insights into the linear model. If the residuals are randomly scattered around zero, it indicates that the model fits the data well. However, if there are deviations or patterns in the residuals, this may suggest that other factors are affecting the data that the model does not account for.
    
### Visualization i was interested in. 


```{r}
incidents_by_location <- nypd_cases_1 %>%
    count(Location_Of_Incident)

# Create the bar plot
ggplot(incidents_by_location, aes(x = Location_Of_Incident, y = n)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    labs(
        title = "Number of Incidents by Location of Incident",
        x = "Location of Incident",
        y = "Number of Incidents"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


### Visualization


```{r}
nypd_cases_1 <- nypd_cases_1 %>%
    mutate(Year = year(Occur_Date))

incidents_per_year_location <- nypd_cases_1 %>%
    group_by(Year, Location_Of_Incident) %>%
    summarize(Incident_Count = n(), .groups = 'drop')

ggplot(incidents_per_year_location, aes(x = Year, y = Incident_Count, color = Location_Of_Incident)) +
    geom_line(linewidth = 1) +
    geom_point(size = 2) +
    labs(
        title = "Number of Incidents per Year by Location",
        x = "Year",
        y = "Number of Incidents",
        color = "Location of Incident"
    ) +
    theme_minimal() +
    scale_x_continuous(breaks = seq(min(incidents_per_year_location$Year), max(incidents_per_year_location$Year), by = 1)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    theme(legend.position = "bottom")
```


#### I wanted to see what the visualization would look like with the number of cases made based on location and in the years followed. since the last visualization was with all locations added.

